The dictionary for characters of smooth group schemes

### 2016-04-19 followup.tex
Deleting extraneous exposition in 4.2; adding references

Let $R$ be a complete discrete valuation ring with maximal ideal $\mathfrak{m}$ and perfect residue field $\Fq$. 
Then $\Spec{R}$ is a henselian trait.
Let $\underline{G}$ be a smooth group scheme over $R$.
Here we shall use  \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg transform.
Let $G$ be the Greenberg transform of $\underline{G}$; then $G$ is a group scheme over $\Fq$ and there is a canonical isomorphism
\[
G(\Fq) = \underline{G}(R).
\]

\begin{proposition}\label{prop:quasicharacters}
With notation as above, suppose $\Fq$ is a finite field.
For every quasicharacter character $\varphi : \underline{G}(S) \to \EEx$ \todo{What is $S$?} there is a Weil sheaf $\cs{L}$ on $G$ such that \[\trFrob{\cs{L}} =  \varphi.\] \end{proposition}

\begin{proof}
By continuity of $\varphi : \underline{G}(S) \to \EE^\times$, there is some $m \in \NN$ and a factorization
\[
\begin{tikzcd}
\underline{G}(S) \arrow{rr}{\varphi} \arrow{rd} && \EEx\\
& \underline{G}(R/\mathfrak{p}^{m+1}) \arrow{ru}[swap]{\varphi_m} 
\end{tikzcd}
\] 
Set $R_m = R/\mathfrak{p}^{m+1}$ and
set $G_m = \Gr_m^{R}(\underline{G})$, the Greenberg transform of $\underline{G}\times_{\Spec{R}}\Spec{R_m}$.
Then $G_m$ is a smooth group scheme over $\Fq$ and $G_m(\Fq) = \underline{G}(R_m)$.
Using Theorem~\ref{thm:geo}, let $\cs{L}_m$ be a geometrization of the character $\varphi_m: G_m(\Fq) \to \EEx$; so
\[
\trFrob{\cs{L}_m} = \varphi_m
\]
on $G_m(\Fq)$.
%
Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_m$.
Let $\cs{L}$ be the  Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along $G \to G_m$. 
Then $\cs{L}$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.3}, such that $\trFrob{\cs{L}} = \varphi$.
\end{proof}

@@

Let $V$ be a finite-dimensional vector space over a finite field $\Fq$ equipped with a symplectic paring $\langle\ ,\ \rangle : V\times V \to Z$, where $Z$ is a one-dimensional vector space over $\Fq$.
Let $V^\sharp$ be the Heisenberg group determined by $(Z, \langle\ ,\ \rangle)$. \todo{We should include a reference defining the Heisenberg group}
Then $V^\sharp$ is a split extension of $V$ by $Z$. \todo{split in what category?}
Let $\Sp(V)$ be the symplectic group determined by the symplectic pairing $\langle\ ,\ \rangle$; this group acts on $V^\sharp$.
The group $\Sp(V)\ltimes V^\sharp$ is called the Jacobi group. 
From the construction above, it is clear that the Jacobi group may be viewed as the $\Fq$-points of an algebraic group over $\Fq$; we will refer to that algebraic group as the Jacobi group.

Let $\psi : Z \to \EEx$ be an additive character and let $\omega_\psi$ be the Heisenberg representation on $V^\sharp$ with central character $\psi$. 
The Heisenberg representation determines a representation $\pi_{\psi}$ of $\Sp(V)$ with the same representation space as $\omega_\psi$ and with the defining property: for each $g\in \Sp(V)$, $\pi_\psi(g)$ determines an isomorphism of representations $\omega_\psi^g \to \omega_\psi$.
Let $W_\psi = \pi_\psi \ltimes \omega_\psi$ be the representation of the Jacobi group $\Sp(V)\ltimes V^\sharp$ given by $\omega_\psi$ and $\pi_\psi$.

The character $\trace(W_\psi) : \Sp(V)\ltimes V^\sharp \to \EE^\times$ of $W_\psi$ is studied in \cite{howe:73a}, where is it shown that $\trace(W_\psi)$ is determined by its restriction to $\Sp(V)$ and by $\psi$.  Moreover, $\trace(W_\psi)(g,v,z) =0$ unless $v\in \ker(g-1)$, in which case $(g,v,z)$ lies in a conjugacy class intersecting $\Sp(V) \ltimes Z$. 
%
The group $\Sp(V)$ is naturally stratified into the conjugation-stable subvarieties $S_r = \{ g\in \Sp(V) \tq \dim \ker (g-1) = r \}$. 
The big stratum, $S_0$, is dense in $\Sp(V)$.
%the boundary of $S_0$ is $\cup_{0< r\leq \dim V} S_r$. 
 From \cite{howe:73a}*{Prop. 2} we learn that if $g\in S_r$ then $q^{-r/2} \trace(W_\psi)(g)$ has complex norm $1$. \todo{We should say something about $\CC$ vs $\EE$}

It is one of the main results of \cite{gurevich-hadani:07a} \todo{Can we give a specific reference here?} (see also \cite{gurevich-hadani:11a}*{Thm. 4.5}) that there is a Weil sheaf complex $\cs{K}_\psi$ on $\Sp(V)\ltimes V^\sharp$ such that 
\begin{equation}
\trFrob{\cs{K}_\psi} = \trace(W_\psi).
\end{equation}
%
Since $\cs{K}_\psi$ is an object in Deligne's category $D^b_c(\Sp(V)\ltimes V^\sharp,\EE)$, the left hand side of this equality must be interpreted accordingly.  Although only the existence of $\cs{K}_\psi$ will be needed below, we recall its construction briefly.

Fix an isomorphism $\CC \approx \EE$ so that we may mix classical and geometric facts. 
Observe that $S_0$ is the domain of the Cayley transform $\cay : g \mapsto (g-1)^{-1}(g+1)$ on $\Sp(V)$.
If $(g,v,z) \in S_0\times V\times Z$ then
\begin{equation}\label{Cayley}
\trace(W_\psi)(g,v,z) =  \frac{\gamma_\psi^{\dim V}}{q^{\dim V/2} } \sigma(\det(1+\cay(g)))\ \psi\left(\langle \cay(g)v,v\rangle/4 \right) \psi(z),
\end{equation}
 where $\gamma_\psi$ is the Gauss sum for $\psi$ and $\sigma : \Fq^\times \to \EE^\times$ is the quadratic character of $\Fq^\times$.
An adaptation of the non-normalized kernels and the normalization coefficients \cite{gurevich-hadani:07a}*{Appendix B.1} allows for the geometrization of \eqref{Cayley} with a $\Sp(V)\ltimes V^\sharp$-equivariant irreducible Weil sheaf $\cs{K}_0$ (in fact, a Weil local system) on $S_0 \times V^\sharp$ such that 
\[
\trFrob{\cs{K}_0(\frac{\dim V}{2})} = \trace(W_\psi)\vert_{S_0\times V^\sharp},
\]
where $\cs{K}_0(\frac{\dim V}{2})$ is the Tate twist of $\cs{K}$.
%
We set $\cs{K}_\psi \ceq j_{!*} \cs{K}_0[\dim V]$, the perverse extension of $\cs{K}_0 [\dim V]$ along $j : S_0 \times V^\sharp \hookrightarrow \Sp(V) \ltimes V^\sharp$.

>>>>

Let $R$ be a complete discrete valuation ring with maximal ideal $\mathfrak{m}$ and perfect residue field $\Fq$. 
Let $\underline{G}$ be a smooth group scheme over $R$.
Here we shall use  \cite{bertapelle-gonzales:Greenberg} for the definition and fundamental properties of the Greenberg transform.
Let $G$ be the Greenberg transform of $\underline{G}$; then $G$ is a group scheme over $\Fq$ and there is a canonical isomorphism
\[
G(\Fq) = \underline{G}(R).
\]

\begin{proposition}\label{prop:quasicharacters}
With notation as above, suppose $\Fq$ is a finite field.
For every quasicharacter character $\varphi : \underline{G}(R) \to \EEx$ there is a Weil sheaf $\cs{L}$ on $G$ such that \[\trFrob{\cs{L}} =  \varphi.\] \end{proposition}

\begin{proof}
By continuity of $\varphi : \underline{G}(R) \to \EE^\times$, there is some $m \in \NN$ and a factorization
\[
\begin{tikzcd}
\underline{G}(S) \arrow{rr}{\varphi} \arrow{rd} && \EEx\\
& \underline{G}(R/\mathfrak{p}^{m+1}) \arrow{ru}[swap]{\varphi_m} 
\end{tikzcd}
\] 
Set $R_m = R/\mathfrak{p}^{m+1}$ and
set $G_m = \Gr_m^{R}(\underline{G})$, the Greenberg transform of $\underline{G}\times_{\Spec{R}}\Spec{R_m}$.
Then $G_m$ is a smooth group scheme over $\Fq$ and $G_m(\Fq) = \underline{G}(R_m)$.
Using Theorem~\ref{thm:geo}, let $\cs{L}_m$ be a geometrization of the character $\varphi_m: G_m(\Fq) \to \EEx$; so
\[
\trFrob{\cs{L}_m} = \varphi_m
\]
on $G_m(\Fq)$.
%
Recall that the full Greenberg transform $G \ceq \Gr^{R}(\underline{G})$ is a group scheme over $\Fq$ such that $G(\Fq) = \underline{G}(R)$; it comes equipped with a morphism $G \to G_m$.
Let $\cs{L}$ be the  Weil sheaf on $G$ obtained from $\cs{L}_m$ by pullback along $G \to G_m$. 
Then $\cs{L}$ is a quasicharacter sheaf on $G$, in the sense of \cite{cunningham-roe:13a}*{Def 4.3}, such that $\trFrob{\cs{L}} = \varphi$.
\end{proof}

@@

Let $V$ be a finite-dimensional vector space over a finite field $\Fq$ equipped with a symplectic paring $\langle\ ,\ \rangle : V\times V \to Z$, where $Z$ is a one-dimensional vector space over $\Fq$.
Let $V^\sharp$ be the Heisenberg group determined by $(Z, \langle\ ,\ \rangle)$ \cite{gurevich-hadani:07a}*{\S 1.1}.
Let $\Sp(V)$ be the symplectic group determined by the symplectic pairing $\langle\ ,\ \rangle$; this group acts on $V^\sharp$.
The group $\Sp(V)\ltimes V^\sharp$ is called the Jacobi group. 
From the construction above, it is clear that the Jacobi group may be viewed as the $\Fq$-points of an algebraic group over $\Fq$; we will refer to that algebraic group as the Jacobi group.

Let $\psi : Z \to \EEx$ be an additive character and let $\omega_\psi$ be the Heisenberg representation on $V^\sharp$ with central character $\psi$ \cite{gurevich-hadani:07a}*{\S 1.1}. 
The Heisenberg representation determines a representation $\pi_{\psi}$ of $\Sp(V)$ with the same representation space as $\omega_\psi$ and with the defining property: for each $g\in \Sp(V)$, $\pi_\psi(g)$ determines an isomorphism of representations $\omega_\psi^g \to \omega_\psi$.
Let $W_\psi = \pi_\psi \ltimes \omega_\psi$ be the Heisenberg-Weil representation of the Jacobi group $\Sp(V)\ltimes V^\sharp$ given by $\omega_\psi$ and $\pi_\psi$ \cite{gurevich-hadani:07a}*{\S 2.2}.

There is a Weil sheaf complex $\cs{K}_\psi$ on $\Sp(V)\ltimes V^\sharp$ \cite{gurevich-hadani:07a}*{Thm 3.2.2.1} (see also \cite{gurevich-hadani:11a}*{Thm. 4.5}) such that 
\begin{equation}
\trFrob{\cs{K}_\psi} = \trace(W_\psi).
\end{equation}
%
Since $\cs{K}_\psi$ is an object in Deligne's category $D^b_c(\Sp(V)\ltimes V^\sharp,\EE)$, the left hand side of this equality must be interpreted accordingly.
